\thetitle

### 2023-04-26 coqimpl-old.tex
Formatting.

  \begin{itemize}
    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
    \begin{minted}{Coq}
      Axiom detectable:
      forall {T} {a} (f: Bᵢ a -> option T),
        {forall x, f x <> None} + {exists x, f x = None}.
    \end{minted} 
    \item Define $"C"$ as the inductive type using this list, i.e.,
    \begin{minted}{Coq}
Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
    \end{minted}
    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
    \begin{itemize}
      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
      \item a partial recursor
      \begin{minted}[escapeinside=@@]{Coq}
prec : forall (R : Set), ..,
  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,
      D -> option R
      \end{minted}
      \item and $n$ (propositional) computational axioms: for all $i$, 
      \begin{minted}{Coq}
prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
      \end{minted}
    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
    \end{itemize}
  \end{itemize}
  Then there exists an \textbf{embedding} from "C" to "D".
  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of\/
  "C" and "prec" of\/ "D" (both acting like identity) such that
  \begin{minted}[escapeinside=@@]{Coq}
    forall c : C, linv (inj c) = some c
  \end{minted}
\end{theorem}
The reason we have this weird looking \ref{prop:detectable-partiality}
is that the construction of "linv"---at one point, using "prec" to
construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
C) → option C}, generally not provable, unless we have
\mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
This \ref{prop:detectable-partiality} is exactly transforming former to
the latter:
recall that, according to the semantic of inductive type,
\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
"linv" on the substructure of "D". This function points out each of the
translation result from "D" to "C" and some of them are "None" because
of failing. For us, we only care about if \textbf{any} of the
substructure of "D" fails to translate---if any substructure fails,
then the whole translation fails and thus we should have "None" as
return.

>>>>

  \begin{itemize}
    \item \textlabel{(Detectable Partiality)}{prop:detectable-partiality} for each pair $x : A_i \vdash B_i(x)$,
    \begin{minted}{Coq}
      Axiom detectable:
      forall {T} {a} (f: Bᵢ a -> option T),
        {forall x, f x <> None} + {exists x, f x = None}.
    \end{minted} 
    \item Define $"C"$ as the inductive type using this list, i.e.,
    \begin{minted}{Coq}
Inductive C : Set := ... | cᵢ : forall (x : Aᵢ), (Bᵢ x -> C) -> C | ...
    \end{minted}
    \item Assume an arbitrary type \mintinline{Coq}{D : Set} with
    \begin{itemize}
      \item $n$ \textbf{functions} \mintinline{Coq}{dᵢ : forall (x : Aᵢ), (Bᵢ x -> D) -> D}  
      \item a partial recursor
      \begin{minted}[escapeinside=@@]{Coq}
prec : forall (R : Set), ..,
  (rᵢ : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)), ..,
      D -> option R
      \end{minted}
      \item and $n$ (propositional) computational axioms: for all $i$, 
      \begin{minted}{Coq}
prec {R} r₁ r₂ .. rₙ (dᵢ aᵢ bᵢ) 
    = rᵢ aᵢ (fun x => prec {R} r₁ r₂ .. rₙ (bᵢ x))
      \end{minted}
    % where \mintinline{Coq}{(lift rᵢ) : (forall (x : Aᵢ), (Bᵢ x -> option R) -> option R)} is defined as expected
    \end{itemize}
  \end{itemize}
  Then there exists an \textbf{embedding} from "C" to "D".
  More concretely, we can have \mintinline{Coq}{inj : C -> D} and 
  \mintinline{Coq}{linv : D -> option C} defined using the eliminator of\/
  "C" and "prec" of\/ "D" (both acting like identity) such that
  \begin{minted}[escapeinside=@@]{Coq}
    forall c : C, linv (inj c) = some c
  \end{minted}
\end{theorem}
The reason we have this weird looking \ref{prop:detectable-partiality}
is that the construction of "linv"---at one point, using "prec" to
construct "linv", we need to prove \\ \mintinline{Coq}{∀ a, (Bᵢ a → option
C) → option C}, generally not provable, unless we have
\mintinline{Coq}{option (Bᵢ a → C)} as an argument instead.
To resolve that, the \ref{prop:detectable-partiality} can help to transform the former to
the latter:
recall that, according to the semantic of inductive type,
\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
"linv" on the substructure of "D". This function points out each of the
translation result from "D" to "C" and some of them are "None" because
of failing. For us, we only care about if \textbf{any} of the
substructure of "D" fails to translate---if any substructure fails,
then the whole translation fails and thus we should have "None" as
return.

### 2023-04-26 lang-design.tex
Formatting.

At one point, using "prec" to
construct "linv", we need to prove \mintinline{Coq}{∀ a, (Bᵢ a → option
C) → option C}. This is generally not provable, unless we have
\mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
"unless we have the Detectable Partiality assumption instead"?}
\EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.}

This \ref{prop:detectable-partiality} can be used to transform the former to
the latter:
recall that, according to the semantic of inductive type,
\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
"linv" on the substructure of "D". This function points out each of the
translation result from "D" to "C" and some of them are "None" because
of failing. For us, we only care about if \textbf{any} of the
substructure of "D" fails to translate---if any substructure fails,
then the whole translation fails and thus we should have "None" as
return.

@@

We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
\begin{minted}{Coq}
  pdelim : ∀ (P : D -> Set), .., 
  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
\end{minted}
To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017} and dependent eliminator is unavoidable.

\noindent
As expected, "tm_prect_STLC" is axiomatized along with
four equalities describing its computational behaviors, one for each constructor.
For instance, the equality for constructor "tm_abs" is as follows:

>>>>

At one point, using "prec" to
construct "linv", we need to prove \\ \mintinline{Coq}{∀ a, (Bᵢ a → option
C) → option C}. This is generally not provable, unless we have \\
\mintinline{Coq}{option (Bᵢ a → C)} instead.\YZ{Don't you really mean
"unless we have the Detectable Partiality assumption instead"?}
\EDJreply{I am telling a more detailed technical point, i.e., how to use detectable partiality to prove linv.} To resolve that, the \ref{prop:detectable-partiality} can be used to transform the former to
the latter:
recall that, according to the semantic of inductive type,
\mintinline{Coq}{(Bᵢ a → option C)} is actually the result of using
"linv" on the substructure of "D". This function points out each of the
translation result from "D" to "C" and some of them are "None" because
of failing. For us, we only care about if \textbf{any} of the
substructure of "D" fails to translate---if any substructure fails,
then the whole translation fails and thus we should have "None" as
return.

@@

We can support partial dependent eliminator as well, but partial dependent eliminator can deduce "prec" and our "prec" is enough to contruct this left inverse for (non-indexed) inductive type. 
\begin{minted}{Coq}
  pdelim : ∀ (P : D -> Set), .., 
  (rᵢ : (∀ (x : Aᵢ) (w : Bᵢ x -> D), (∀ (b : Bᵢ x), option (P (w b))) 
    -> option (P (dᵢ x w)))),  .., ∀ (d : D) -> option (P d)
\end{minted}
To generalize to indexed inductive type, we need to base on indexed W type~\cite{martin1982constructive, morris2009indexed,jashug2017}.

As expected, "tm_prect_STLC" is axiomatized along with
four equalities describing its computational behaviors, one for each constructor.
For instance, the equality for constructor "tm_abs" is as follows:
